Trans-Dimensional Bayesian Inference for Gravitational Lens Substructures

### 2015-09-03 Papers/ms.tex
slight wording change
Fixed empirical measure equations
Forgot revisions

\section{{\revisions``Easy''} Simulated Data}
To demonstrate the method, we generated a simulated dataset {\revisions where
the lens was an SIE$+\gamma$ profile with a single additional substructure}.
The image is shown in Figure~\ref{fig:image1}, and
consists of 100 $\times$ 100 pixels. The point spread function (PSF) was compact and was
defined on a 5 $\times$ 5 grid. The data was created by firing only one ray
per pixel, and the inference was also carried out under the same level of
approximation. Since the model assumptions are all correct for this image, the
demonstration here is purely to illustrate the computational tractability of
the problem, and the kind of outputs the method can produce.
{\revisions
The SIE$+\gamma$ parameters were
$\{b, q, x_c^{\rm SIE}, y_c^{\rm SIE}, \theta, \gamma, \theta_\gamma\}
= \{5.25, 0.323, 0.191, -0.341, 0.833, 0.0322, 0.497\}$, where the length
units in the image plane are arbitrary, and rotation angles are measured in
radians. The single substructure is located at $(-2.57, 2.20)$, close to the
upper left image. Its mass is 2.50 units, using the convention where the
critical density is 1. For comparison, the SIE mass within its critical
ellipse is $\pi b^2 = 86.6$ units.
}

@@

The posterior expected value of the empirical measure is:
\begin{eqnarray}
\int p(\boldsymbol{\theta} | \boldsymbol{D})
\sum_{i=1}^{N_{\rm lens}}
\delta^2\left(x - x_i^{\rm lens}, y_i^{\rm lens} - y_i^{\rm lens}\right)
\, d\boldsymbol{\theta}
\end{eqnarray}
where $\boldsymbol{\theta}$ denotes all parameters and hyperparameters
(including the $N$s) and ``$\int \,d\boldsymbol{\theta}$'' is an integral and summation over the entire parameter space.

Since we can approximate posterior expectations using Monte Carlo, we can
obtain the expected value of the empirical measure using:
\begin{eqnarray}
\frac{1}{n}
\sum_{k=1}^n
\sum_{i=1}^{N_{\rm lens}}
\delta^2\left(x - x_i^{\rm lens}, y_i^{\rm lens} - y_i^{\rm lens}\right)
\end{eqnarray}
where $n$ is the number of posterior samples. The resulting function is
an ``image'' with a point mass wherever a substructure occurred. For
visualisation purposes the image can be blurred, or calculated at a lower
resolution by replacing the Dirac-delta function with a discrete version
which returns a nonzero constant
if a substructure appears in a pixel or zero otherwise.

>>>>

\section{{\revisions``Easy''} Simulated Data}
To demonstrate the method, we generated a simulated dataset {\revisions where
the lens was an SIE$+\gamma$ profile with a single additional substructure}.
The image is shown in Figure~\ref{fig:image1}, and
consists of 100 $\times$ 100 pixels. The point spread function (PSF) was compact and was
defined on a 5 $\times$ 5 grid. The data was created by firing only one ray
per pixel, and the inference was also carried out under the same level of
approximation. Since the model assumptions are all correct for this image, the
demonstration here is purely to illustrate the computational tractability of
the problem, and the kind of outputs the method can produce.
{\revisions
The SIE$+\gamma$ parameters were
$\{b, q, x_c^{\rm SIE}, y_c^{\rm SIE}, \theta, \gamma, \theta_\gamma\}
= \{5.25, 0.323, 0.191, -0.341, 0.833, 0.0322, 0.497\}$, where the length
units in the image plane are arbitrary, and rotation angles are measured in
radians. The single substructure is located at $(-2.57, 2.20)$, about halfway
between the center and the upper left image.
Its mass is 2.50 units, using the convention where the
critical density is 1. For comparison, the SIE mass within its critical
ellipse is $\pi b^2 = 86.6$ units.
}

@@

The posterior expected value of the empirical measure is:
{\revisions
\begin{eqnarray}
\int p(\boldsymbol{\theta} | \boldsymbol{D})
\sum_{i=1}^{N_{\rm lens}}
\delta^2\left(x - x_i^{\rm lens}, y - y_i^{\rm lens}\right)
\, d\boldsymbol{\theta}
\end{eqnarray}
}
where $\boldsymbol{\theta}$ denotes all parameters and hyperparameters
(including the $N$s) and ``$\int \,d\boldsymbol{\theta}$'' is an integral and summation over the entire parameter space.

Since we can approximate posterior expectations using Monte Carlo, we can
obtain the expected value of the empirical measure using:
{\revisions
\begin{eqnarray}
\frac{1}{n}
\sum_{k=1}^n
\sum_{i=1}^{N_{\rm lens}^k}
\delta^2\left(x - x_{ik}^{\rm lens}, y - y_{ik}^{\rm lens}\right)
\end{eqnarray}
}
where $n$ is the number of posterior samples. The resulting function is
an ``image'' with a point mass wherever a substructure occurred. For
visualisation purposes the image can be blurred, or calculated at a lower
resolution by replacing the Dirac-delta function with a discrete version
which returns a nonzero constant
if a substructure appears in a pixel or zero otherwise.
